Visualizations for the manuscript

### 2022-05-02 metanalysis_visualization/manuscript_Figures.Rmd
tweaking figures

```{r}
# Call function from another file : bootstrap.R
bootstrap_size <- 10000
binom_proportion <- function(n, prob, size) {
  (rbinom(n, size, prob) / size) * 100
}

@@

```{r}
annotation_function <- function(df, ...) {
  z <- unique(df[["observed_diff"]]) / sd(df[["periodduring_pandemic:time"]])
  pvalue_coeff_interaction <- round(pnorm(abs(z), lower.tail = FALSE) * 2, 3)
  paste0("Coeffs differences: ",
         round(unique(df[["observed_diff"]]), 2),
         " (pvalue: ",
         format.pval(pvalue_coeff_interaction),
         ")")
}
```

@@

```{r}
df_coefficients_boot <- bootstrapped_coefficients_all %>% 
  left_join(coefficients_df, by = "siteid")
names_site_id <- unique(df_coefficients_boot$siteid)
plot_annotations <- df_coefficients_boot %>% 
  group_by(siteid) %>% 
  group_map( annotation_function ) %>% 
  set_names(names_site_id) %>% 
  unlist()
plot_annotations_df <- data.frame(
  siteid = names(plot_annotations), 
  label = plot_annotations
)
```

@@

```{r}
sexInfo_per_country <- table1_all %>%
  filter( Category == "Sex") %>%
  mutate ( counts_before = as.numeric(sapply(strsplit( as.character(before_pandemic), " "), '[', 1)), 
           counts_during = as.numeric(sapply(strsplit( as.character(during_pandemic), " "), '[', 1))) %>%
  select( name, siteid, counts_before, counts_during ) %>%
  inner_join( country_map ) %>%
  group_by( name, country ) %>%
  summarise( before = sum( counts_before), 
             during = sum( counts_during )) %>%
  pivot_longer(names_to = "period", cols=c(before, during)) 
sexInfo_all <- table1_all %>%
  filter( Category == "Sex") %>%
  mutate ( counts_before = as.numeric(sapply(strsplit( as.character(before_pandemic), " "), '[', 1)), 
           counts_during = as.numeric(sapply(strsplit( as.character(during_pandemic), " "), '[', 1))) %>%
  select( name, siteid, counts_before, counts_during ) %>%
  inner_join( country_map ) %>%
  group_by( name ) %>%
  summarise( before = sum( counts_before), 
             during = sum( counts_during )) %>%
  pivot_longer(names_to = "period", cols=c(before, during)) %>%
  mutate( country = 'ALL' ) %>%
  select( name, country, period, value ) %>%
  rbind( sexInfo_per_country)
sexInfo_all %>%
  ggplot( aes( x= name, y = value, fill = period )) +
  geom_bar( stat = "identity", position = "dodge") +
  facet_wrap("country", scales = "free_y") + 
  labs(x ="", y = "number of patients") +
  theme(axis.text.x=element_text(hjust=1, size = 8), 
        strip.text = element_text(size = 10, face = "bold"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA))+ 
  scale_fill_manual("Pandemic \n period", values = c("before" = "grey", "during" = "orange"))

```{r}
patients_hospitalized_agg_sex_total <- patients_hospitalized_agg_sex_perc_all %>%
  select( - percentage ) %>%
  group_by( time_p, sex ) %>%
  summarise( total = sum( count ), 
             count_sex = sum( count_sex ), 
             percentage = 100* count_sex / total, 
             siteid = "ALL")

patients_hospitalized_agg_sex_total %>%
  ggplot( aes( x= as.Date( time_p ), y = percentage, group = sex )) +
  geom_line( aes( linetype = sex , color = sex)) +
  geom_point( aes( shape = sex, color = sex )) +
  scale_colour_manual(values=c("#FF6600","#339999")) +
  scale_fill_manual(values=c("#FF6600","#339999"))+
  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed")+
  labs(x ="", y = "percentage of patients (%)")+
  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA))

@@

```{r }
bootstrapped_sex <- patients_hospitalized_agg_sex_total %>%
  group_by(sex) %>% 
  # group_by(plotsLabel, sex) %>% 
  mutate(month_year = format(time_p, "%m"),
         time = seq_along(time_p))  %>% 
 group_map(~ bootstrapping_slopes(.x,
                                  group_name = .y,
                                  bootstrap_size = bootstrap_size,
                                  count = "count_sex", 
                                  percentage = "percentage"))

@@

## Using only male, but female and male necessarily yield equal p-values
plot_annotations_sex_df <- annotation_function(
  df_coefficients_boot_sex[df_coefficients_boot_sex$sex == "female", ]
) %>% data.frame(label = ., 
                 plotsLabel = "Female")
```

```{r bootstrap_plot}
ggplot(bootstrapped_fitted_sex, aes(x = time_p, y = predicted, group = interaction(iteration, period, sex))) +
  # facet_wrap("plotsLabel", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
  geom_line(aes(color = interaction(period, sex)),
            alpha = 0.03) +
  geom_line(data = quantile_bootstrapped_fitted_sex,
            aes(y = value, group = interaction(quantiles, period, sex)),
            linetype = "dashed") +
  geom_line(data = quantile_bootstrapped_fitted_sex[quantile_bootstrapped_fitted_sex$quantiles == "50%", ],
            aes(y = value, group = interaction(period, sex))) +
  geom_point(data = patients_hospitalized_agg_sex_total,
             aes(y = percentage, group=sex, color = interaction(period, sex)),
             alpha = 1) +
  scale_color_brewer(palette = "Paired") +
  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
  # scale_colour_manual(values = c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C")) +
  guides(color = guide_legend(override.aes = list(alpha = 1))) +
  geom_label(data = plot_annotations_sex_df,
             aes(label=label),
            x = -Inf,
            y = +Inf,
            hjust = 0,
            vjust = 1,
            size = 4,
            inherit.aes = FALSE) +
  labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
  subtitle = "Parametric bootstrap estimation: (10000 iterations)",
       x = "Calendar dates",
       y = "Percentage",
       color = "Period") +
  theme(strip.text = element_text(size = 10),
        axis.text.x=element_text(angle = 45, hjust=1, size = 8),
        axis.text.y = element_text(size = 8),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        title = element_text(size = 14),
  plot.subtitle = element_text(size = 12, face = "italic"), 
  legend.position = "bottom")
ggsave("figure4_ITS_sex.png", height = 20, width = 20, units = "cm")
```

@@

plot_annotations_disorder <- df_coefficients_boot_disorder %>% 
  group_by(disorder_group) %>% 
  group_map( annotation_function ) %>% 
  set_names(names_disorder) %>% 
  unlist()
plot_annotations_disorder_df <- data.frame(
  disorder_group = names(plot_annotations_disorder), 
  label = plot_annotations_disorder
)
```

@@

### By site
```{r}
sites <- unique( patient_count_psy_period_psy_all$siteid )
source("../metanalysis_visualization/time_series_analysis_plotting.R")
p <- list()

@@

>>>>

```{r}
# Call function from another file : bootstrap.R
bootstrap_size <- 1000
binom_proportion <- function(n, prob, size) {
  (rbinom(n, size, prob) / size) * 100
}

@@

```{r}
bootstrap_pvalue_calculation <- function(df, ...) {
  z <- unique(df[["observed_diff"]]) / sd(df[["periodduring_pandemic:time"]])
  pvalue_coeff_interaction <- round(pnorm(abs(z), lower.tail = FALSE) * 2, 3)
  list("pvalue" = pvalue_coeff_interaction, 
       "observed_diff" = round(unique(df[["observed_diff"]]), 2)
  ) 
}

@@

```{r}
df_coefficients_boot <- bootstrapped_coefficients_all %>% 
  left_join(coefficients_df, by = "siteid")
names_site_id <- unique(df_coefficients_boot$siteid)

plot_annotations <- df_coefficients_boot %>% 
  group_by(siteid) %>% 
  group_map( bootstrap_pvalue_calculation ) %>% 
  set_names(names_site_id) %>% 
  lapply(FUN = function(vec) {
    paste0("Coeffs differences: ",
           vec$observed_diff,
           " (pvalue: ",
           format.pval(vec$pvalue, 
                       eps = 10**-5),
           ")")
  }) %>% 
  unlist()

plot_annotations_df <- data.frame(
  siteid = names(plot_annotations), 
  label = plot_annotations
)
```

@@

```{r, include=FALSE}
patients_hospitalized_agg_sex_total <- patients_hospitalized_agg_sex_perc_all %>%
  select( - percentage ) %>%
  group_by( time_p, sex ) %>%
  summarise( total = sum( count ), 
             count_sex = sum( count_sex ), 
             percentage = 100* count_sex / total, 
             siteid = "ALL")
```

@@

```{r }
bootstrapped_sex <- patients_hospitalized_agg_sex_total %>%
  group_by(sex) %>% 
  # group_by(plotsLabel, sex) %>% 
  mutate(month_year = format(time_p, "%m"),
         time = seq_along(time_p))  %>% 
 group_map(~ bootstrapping_slopes(.x,
                                  group_name = .y,
                                  bootstrap_size = 10000,
                                  count = "count_sex", 
                                  percentage = "percentage"))

@@

## Using only male, but female and male necessarily yield equal p-values

pvalue_sex_slope_diff <- bootstrap_pvalue_calculation(  df_coefficients_boot_sex[df_coefficients_boot_sex$sex == "female", ]
)

plot_annotations_sex_df <- paste0("Female/male proportion coefficient differences: ",
                                  pvalue_sex_slope_diff$observed_diff,
                                  " (pvalue: ",
                                  format.pval(pvalue_sex_slope_diff$pvalue, 
                                              eps = 10**-5),")") %>%
  data.frame(label = ., 
             plotsLabel = "Female")
```

```{r bootstrap_plot}
ggplot(bootstrapped_fitted_sex, aes(x = time_p, y = predicted/100, group = interaction(iteration, period, sex))) +
  geom_line(aes(color = interaction(period, sex)),
            alpha = 0.003) +
  geom_line(data = quantile_bootstrapped_fitted_sex,
            aes(y = value/100, group = interaction(quantiles, period, sex)),
            linetype = "dashed") +
  geom_line(data = quantile_bootstrapped_fitted_sex[quantile_bootstrapped_fitted_sex$quantiles == "50%", ],
            aes(y = value/100, group = interaction(period, sex))) +
  geom_point(data = patients_hospitalized_agg_sex_total,
             aes(y = percentage/100, group=sex),
             alpha = 1) +
  # scale_color_brewer(palette = "Paired") +
  scale_color_manual(labels = c("Female", "Male", "", ""),
                     values = c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C")) +
  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
  guides(colour = guide_legend(
    override.aes = list(alpha = 1,
                        colour = c('#1F78B4', '#33A02C', NA, NA))) 
  ) +
  geom_label(data = plot_annotations_sex_df,
             aes(label = label),
             x = -Inf,
             y = +Inf,
             hjust = 0,
             vjust = 1,
             size = 4,
             inherit.aes = FALSE) +
  labs(title = "Interrupted Time Series Analysis: Patients sex proportion",
       subtitle = "Parametric bootstrap estimation: (10000 iterations)",
       x = "Calendar dates",
       y = "Sex proportion",
       color = "Sex") +
  theme(strip.text = element_text(size = 10),
        axis.text.x=element_text(angle = 45, hjust=1, size = 8),
        axis.text.y = element_text(size = 8),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        title = element_text(size = 14),
  plot.subtitle = element_text(size = 12, face = "italic"))
ggsave("figure4_ITS_sex.png", height = 20, width = 20, units = "cm")
```

@@

plot_annotations_disorder <- df_coefficients_boot_disorder %>% 
  group_by(disorder_group) %>% 
  group_map( bootstrap_pvalue_calculation ) %>%
  set_names(names_disorder) %>% 
  lapply(FUN = function(vec) {
    paste0("Coeffs differences: ",
           vec$observed_diff,
           " (pvalue: ",
           format.pval(vec$pvalue, 
                       eps = 10**-5),
           ")")
  }) %>% 
  unlist()
plot_annotations_disorder_df <- data.frame(
  disorder_group = names(plot_annotations_disorder), 
  label = plot_annotations_disorder
)
```

@@

@@

# Unused figures

```{r, include = FALSE}
sexInfo_per_country <- table1_all %>%
  filter( Category == "Sex") %>%
  mutate ( counts_before = as.numeric(sapply(strsplit( as.character(before_pandemic), " "), '[', 1)), 
           counts_during = as.numeric(sapply(strsplit( as.character(during_pandemic), " "), '[', 1))) %>%
  select( name, siteid, counts_before, counts_during ) %>%
  inner_join( country_map ) %>%
  group_by( name, country ) %>%
  summarise( before = sum( counts_before), 
             during = sum( counts_during )) %>%
  pivot_longer(names_to = "period", cols=c(before, during)) 
sexInfo_all <- table1_all %>%
  filter( Category == "Sex") %>%
  mutate ( counts_before = as.numeric(sapply(strsplit( as.character(before_pandemic), " "), '[', 1)), 
           counts_during = as.numeric(sapply(strsplit( as.character(during_pandemic), " "), '[', 1))) %>%
  select( name, siteid, counts_before, counts_during ) %>%
  inner_join( country_map ) %>%
  group_by( name ) %>%
  summarise( before = sum( counts_before), 
             during = sum( counts_during )) %>%
  pivot_longer(names_to = "period", cols=c(before, during)) %>%
  mutate( country = 'ALL' ) %>%
  select( name, country, period, value ) %>%
  rbind( sexInfo_per_country)
sexInfo_all %>%
  ggplot( aes( x= name, y = value, fill = period )) +
  geom_bar( stat = "identity", position = "dodge") +
  facet_wrap("country", scales = "free_y") + 
  labs(x ="", y = "number of patients") +
  theme(axis.text.x=element_text(hjust=1, size = 8), 
        strip.text = element_text(size = 10, face = "bold"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA))+ 
  scale_fill_manual("Pandemic \n period", values = c("before" = "grey", "during" = "orange"))

```{r, include=FALSE}
patients_hospitalized_agg_sex_total %>%
  ggplot( aes( x= as.Date( time_p ), y = percentage, group = sex )) +
  geom_line( aes( linetype = sex , color = sex)) +
  geom_point( aes( shape = sex, color = sex )) +
  scale_colour_manual(values=c("#FF6600","#339999")) +
  scale_fill_manual(values=c("#FF6600","#339999"))+
  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed")+
  labs(x ="", y = "percentage of patients (%)")+
  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill = NA))

### By site
```{r, include=FALSE}
sites <- unique( patient_count_psy_period_psy_all$siteid )
source("../metanalysis_visualization/time_series_analysis_plotting.R")
p <- list()
